Lessons Learned about JWST Transiting Planet Time Series from Ground-based Studies of the NIRCam Detectors

### 2020-03-12 ms.tex
more studies of read noise covariance matrix and optimal weighting

However, if one generalizes the weighted average to a covariance matrix \citep{schmelling1995averagingCorrelatedData}, the weights become,
\begin{equation}
w_j = \sum_{i=1}^{m} \left( C^{-1}\right)_{ij} P_j,\label{eq:covarianceWeighted}
\end{equation}
where $C^{-1}$ is the inverse of the covariance matrix between pixels $i$ and $j$ and $m$ is the number of pixels in the summation.
Simply variance weighting the spectra by the read and photon noise does not provide heavy enough weights.
If we assume the read noise is perfectly correlated along the spectral direction and uncorrelated along the spatial direction, then equation \ref{eq:covarianceWeighted} becomes
\begin{equation}
w_j = \frac{P_j}{N_{ph} + N_{px} \sigma_R^2},
\end{equation}
where $N_{px}$ is the number of spectral pixels that will be binned and $\sigma_R$ is the read noise per pixel.
Thus, for large values of $N_{px}$ (up to 512 for an amplifier), the $\sigma_R \approx 14e^-$ of read noise per pixel, can compound to 100,400 e$^e-$ and the weighting approaches
\begin{equation}
w_j \propto P_j.
\end{equation}
Profile weighting should therefore lower the error on wavelength-binned time series.
For the grism aperture experiment, we find that profile weighting gives a measured read noise of 355 ppm.

\subsubsection{Empirical Covariance Matrix}

\begin{figure}[!hbtp]
\centering
\includegraphics[width=.7\columnwidth]{spec_spatial_cov.pdf}
\caption{The read noise is strongly correlated in the spectral direction (diagnoal boxes) and somewhat correlated in the spatial direction (off diagonal region).}\label{fig:covSpeSpatial}
\end{figure}

>>>>

However, if one generalizes the weighted average to a covariance matrix \citep{schmelling1995averagingCorrelatedData}, the weights become,
\begin{equation}
w_j = \sum_{i=1}^{m} \left( C^{-1}\right)_{ij} P_j,\label{eq:covarianceWeighted}
\end{equation}
where $C^{-1}$ is the inverse of the covariance matrix between pixels $i$ and $j$ and $m$ is the number of pixels in the summation.
Simply variance weighting the spectra by the read and photon noise does not provide heavy enough weights.
If we assume the read noise is perfectly correlated along the spectral direction and uncorrelated along the spatial direction, then equation \ref{eq:covarianceWeighted} becomes
\begin{equation}
w_j = \frac{P_j}{N_{ph} + N_{px} \sigma_R^2},
\end{equation}
where $N_{px}$ is the number of spectral pixels that will be binned and $\sigma_R$ is the read noise per pixel.
Thus, for large values of $N_{px}$ (up to 512 for an amplifier), the $\sigma_R \approx 14e^-$ of read noise per pixel, can compound to 100,400 e$^-$ and the weighting approaches
\begin{equation}
w_j \propto P_j.
\end{equation}
Profile weighting should therefore lower the error on wavelength-binned time series.
For the grism aperture experiment, we find that profile weighting gives a measured read noise of 355 ppm.
Thus, profile weighting still does not reduce the standard deviation over an extraction aperture that is just 1 pixel wide at the peak grism flux.
We show the real covariance matrix in Section \ref{sec:empCovarianceMatrix}.

\subsubsection{Empirical Covariance Matrix}\label{sec:empCovarianceMatrix}
The ideal weighting of the spectrum from Equation \ref{eq:covarianceWeighted} depends on the covariance matrix $C_{ij}$.
We use CDS (correlated double sampling) read pairs of OTIS long dark data (108 frames) with reference pixel corrections to estimate $C_{ij}$.
We consider a wavelength bin that is 100 pixels in the dispersion direction (0.1$\mu$m) and 14 spatial spatial pixels.
Figure \ref{fig:covSpecSpatial} shows the covariance matrix of this wavelength bin, taken from the median of 20 regions of the detector.
The diagonal elements are consistent with NIRCam detector performance with $\sigma^2 \approx 13^2 (e^-)^2=169 (e^-)^2$.
It is clear that 1/f noise causes covariance most prominently along the spectral direction (the fast-read direction), with a covariance nearly uniformly at 36~$(e^-)^2$.
Surprisingly, there is also a covariance between spatial pixels (the slow read direction) of $\sim$16 $(e^-)^2$ that falls off to $\sim 10 (e^-)^2$.

\begin{figure}[!hbtp]
\centering
\includegraphics[width=.7\columnwidth]{spec_spatial_cov.pdf}
\caption{The read noise is strongly correlated in the spectral direction (diagnoal boxes) and somewhat correlated in the spatial direction (off diagonal region).}\label{fig:covSpecSpatial}
\end{figure}
